---
title: "Low level implementation details"
author: "Rich FitzJohn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Low level implementation details}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

These are some implementation details that may be of interest if
developing alternative (non-hiredis based) drivers.

Create a connection:
``` {r }
con <- redux::redis_connection(redux::redis_config())
con
```

See `?redux::redis_config` for details on specifying hosts,
ports, servers, etc.  Importantly, _socket connections_ can be
used, which can be considerably faster if you are connecting to a
local Redis instance and have socket connections enabled.

The connection object provides several functions for interfacing
with Redis:

* `config`: return the configuration used to create the connection
* `reconnect`: Reconnect a disconnected client, using the same
  options as used to create the connection.
* `command`: Run a redis command.  The format of the argument is
  given below.
* `pipeline`: Run a set of redis commands, as described below, in a
  single roundtrip with the server.  No logic is possible here, but
  this can be *much* faster especially over slow connections.
* `subscribe`: Support for becoming a blocking subscribe client.

## Commands

The simplest command is a character vector, starting with a Redis
command, e.g.:
``` {r }
con$command(c("SET", "foo", 1))
```

However, if we wanted to set `foo` to be an R object (e.g. `1:10`),
then we need to *serialise* the object.  To do that we use `raw`
vectors and the command would become:
``` {r }
con$command(list("SET", "foo", serialize(1:10, NULL)))
```

and to retrieve it:
``` {r }
unserialize(con$command(c("GET", "foo")))
```

Elements of a list command can be:

* Character vectors of any length
* Integer vectors of any length (converted to character)
* Logical vectors of any length (converted to integer then to character)
* A list of raw vectors (note this generates a nested list)

`NULL` values in the list will be skipped over.

## Pipelining

See [the redis documentation](http://redis.io/topics/pipelining)
for background information about pipelining.  In short, evaluating
``` {r }
con$command(c("INCR", "X"))
con$command(c("INCR", "X"))
```

will result in two round trips:

```
Client: INCR X
Server: 1
Client: INCR X
Server: 2
```

These commands can be *pipelined* together into a single request so
that the interaction looks like:

```
Client: INCR X
Client: INCR X
Server: 1
Server: 2
```

To do this, the `pipeline` function in the redis object accepts
multiple Redis commands as a list:
``` {r }
con$pipeline(list(
  c("INCR", "X"),
  c("INCR", "X")
))
```

(if these arguments are named, then the output will have the same
names).

Note the [warnings about
pipeline](http://redis.io/topics/pipelining#redis-pipelining) in
the official Redis documentation - sending so many commands (e.g.,
>10k) memory use on the server can be negatively affected.

## Subscriptions

Subscriptions, really should be done with the wrapper, which is
exposed by the `subscribe` method of a `redis_api` object (e.g,
`redux`).  The brave are welcome to use this low-level interface
should the need arise.  The `subscribe` function takes arguments:

* `channel`: a vector of one or more channels to listen on
* `pattern`: a logical indicating if the `channel` is a pattern
* `callback`: a function described below
* `envir`: environment to evaluate the function in, by default the
  parent frame

The callback function must take a single argument; this will be the
received message with named elements `type` (which will be
message), `channel` (the name of the channel) and `value` (the
message contents).  If `pattern` was `TRUE`, then an additional
element `pattern` will be present (see the Redis docs).  The
callback must return `TRUE` or `FALSE`; this indicates if the
client should continue quit (i.e., `TRUE` means return control to
R, `FALSE` means keep going).

Because the `subscribe` function is blocking and returns nothing,
so all data collection needs to happen as a side-effect of the
callback function.

Here's an example that will collect values until it has 10 entries:
``` {r }
callback <- local({
  i <- 1L
  vals <- numeric(10L)
  function(x) {
    vals[[i]] <<- as.numeric(x$value)
    i <<- i + 1L
    i > 10L
  }
})
```
which, given a valid publisher would look like:

```r
con$subscribe("foo", FALSE, callback)
```

This will sit there forever unless something publishes on channel
`foo`.  In a second instance you can run:

```r
con <- redux::redis_connection(redux::redis_config())
res <- sapply(1:11, function(i) con$command(c("PUBLISH", "foo", runif(1))))
```

which will return 10 values of 1 and 1 value of 0 (being the number of clients subscribed to the `foo` channel).

Back in the client R instance, the subscriber has detached, and

```r
environment(callback)$vals
```

will be a vector of 10 random numbers!
