---
title: "Using Redis with redux and RedisAPI"
author: "Rich FitzJohn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Redis with redux and RedisAPI}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

``` {r echo=FALSE,results="hide"}
library(methods)
knitr::opts_chunk$set(error=FALSE)
redux::hiredis()$DEL(c("mykey", "mylist", "mylist2", "rlist"))
set.seed(1)
```

`redux` and `RedisAPI` provide a full interface to the Redis API;
`redux` does the actual interfacing with Redis and `RedisAPI` uses
this to expose all `r length(grep("^[A-Z]",
names(RedisAPI:::redis_api_generator$public_methods)))` Redis
commands as a set of user-friendly R functions that do basic error
checking.

It is possible to build user-friendly applications on top of this,
for example, the built in `rdb` R key-value store,
[`storr`](https://github.com/richfitz/storr) which provides a
content-addressable object store, and
[`rrqueue`](https://github.com/traitecoevo/rrqueue) which
implements a scalable queuing system.

The main entry point for creating a `redis_api` object is the
`hiredis` function:
``` {r }
r <- redux::hiredis()
```

By default, it will connect to a database running on the local
machine (`localhost`, or ip `127.0.0.1`) and port 6379.  The two
arguments that `hiredis` accepts are `host` and `port` if you need
to change these to point at a different location.  Alternatively,
you can set the `REDIS_HOST` and `REDIS_PORT` environment variables
to appropriate values and then use `hiredis()` with no arguments.

The `redis_api` object is an [`R6`](https://github.com/wch/R6)
class with _many_ methods, each corresponding to a different Redis
command.
``` {r eval=FALSE}
r
```
``` {r echo=FALSE}
res <- capture.output(print(r))
res <- c(res[1:6], "    ...", res[(length(res) - 3):(length(res) - 1)])
writeLines(res)
```

For example, `SET` and `GET`:
``` {r }
r$SET("mykey", "mydata") # set the key "mykey" to the value "mydata"
r$GET("mykey")
```

# Serialisation

The value for most arguments must be a string or will be coerced
into one.  So if you want to save an arbitrary R object, you need
to convert it to a string.  The `object_to_bin` and
`object_to_string` functions can help here, serialising the objects
to binary and string represenations.
``` {r }
obj <- RedisAPI::object_to_bin(1:10)
obj
```

or
``` {r }
str <- RedisAPI::object_to_string(1:10)
str
```

The binary serialisation is faster, smaller, and preserves all the
bits of floating point numbers.  The string version might be
preferable where having only strings in the database is wanted.
The binary serialisation is compatible with the same approach used
in `RcppRedis`, though it is never done automatically.

These values can be deserialised:
``` {r }
RedisAPI::bin_to_object(obj)
RedisAPI::string_to_object(str)
```

So:
``` {r }
r$SET("mylist", RedisAPI::object_to_bin(1:10))
r$GET("mylist")
RedisAPI::bin_to_object(r$GET("mylist"))
```

Or:
``` {r }
r$SET("mylist", RedisAPI::object_to_string(1:10))
r$GET("mylist")
RedisAPI::string_to_object(r$GET("mylist"))
```

This gives you all the power of Redis, but you will have to
manually serialise/deserialise all complicated R objects (i.e.,
everything other than logicals, numbers or strings).  Similarly,
you are responsible for type coersion/deserialisation when
retrieving data at the other end.

This is how the `rdb` object is implemented (see below).

However, depending on what you want to achieve, Redis offers
potentially better ways of holding lists using its native data
types.  For example;

``` {r }
r$RPUSH("mylist2", 1:10)
```

(the returned value `10` indicates that the list "mylist2" is 10
elements long).  There are lots of commands for operating on lists:

``` {r }
RedisAPI::redis_commands("list")
```

For example, you can do things like;

* get an element by its index (note tht this uses C-style base-0
indexing for consistency with the `Redis` documentation rather than
R's semantics)
``` {r }
RedisAPI::redis_help("LINDEX")
r$LINDEX("mylist2", 1)
```

* set an element by its index
``` {r }
RedisAPI::redis_help("LSET")
r$LSET("mylist2", 1, "carrot")
```

* get all of a list:
``` {r }
RedisAPI::redis_help("LRANGE")
r$LRANGE("mylist2", 0, -1)
```

* or part of it:
``` {r }
r$LRANGE("mylist2", 0, 2)
```

* pop elements off the front or back
``` {r }
RedisAPI::redis_help("LLEN")
RedisAPI::redis_help("LPOP")
RedisAPI::redis_help("RPOP")
r$LLEN("mylist2")
r$LPOP("mylist2")
r$RPOP("mylist2")
r$LLEN("mylist2")
```

Of course, each element of the list can be an R object if you run
it through `object_to_string`:
``` {r }
r$LPUSH("mylist2", RedisAPI::object_to_string(1:10))
```

but you'll be responsible for converting that back (and detecting
/ knowing that this needs doing)
``` {r }
dat <- r$LRANGE("mylist2", 0, 2)
dat
dat[[1]] <- RedisAPI::string_to_object(dat[[1]])
dat
```

# Example application of an R-object key/value store (`rdb`)

As a simple example, the `RedisAPI` package provides a `rdb` object
showing how one might build an application on top of `RedisAPI` and
`redux`.  To create an `rdb` object
``` {r }
db <- RedisAPI::rdb(redux::hiredis)
```

Newly created databases are empty - they have no keys
``` {r }
db$keys()
```

R objects can be stored against keys, for example:
``` {r }
db$set("mykey", 1:10)
db$keys()
```

Retrieve the value of a key with `$get`:
``` {r }
db$get("mykey")
```

Trying to get a nonexistant key does not throw an error but returns
`NULL`
``` {r }
db$get("no_such_key")
```

That's it.  Arbitrary R objects can be stored in keys, and they
will be returned intact with few exceptions (the exceptions are
things like `rdb` itself which includes an "external pointer"
object which can't be serialised - see `?serialize` for more
information):
``` {r }
db$set("mtcars", mtcars)
identical(db$get("mtcars"), mtcars)
db$set("a_function", sin)
db$get("a_function")(pi / 2) # 1
```

This seems really silly, but is potentially very useful.  There are
file-based key/value systems on CRAN, and this would be another but
backed by a datea source that is distributed.

# Potential applications

Because `RedisAPI` exposes all of Redis, you can roll your own data
structures.

First, a generator object that sets up a new list at `key` within
the database `r`.
``` {r }
rlist <- function(..., key="rlist", r=redux::hiredis()) {
  dat <- vapply(c(...), RedisAPI::object_to_string, character(1))
  r$RPUSH(key, dat)
  ret <- list(r=r, key=key)
  class(ret) <- "rlist"
  ret
}
```

Then some S3 methods that work with this object.  I've only
implemented `length` and `[[`, but `[` would be useful here too as
would `print`.
``` {r }
length.rlist <- function(x) {
  x$r$LLEN(x$key)
}

`[[.rlist` <- function(x, i, ...) {
  RedisAPI::string_to_object(x$r$LINDEX(x$key, i - 1L))
}

`[[<-.rlist` <- function(x, i, value, ...) {
  x$r$LSET(x$key, i - 1L, RedisAPI::object_to_string(value))
  x
}
```

Then we have this weird object we can add things to.
``` {r }
obj <- rlist(1:10)
length(obj) # 10
obj[[3]]
obj[[3]] <- "an element"
obj[[3]]
```

The object has reference semantics so that assignment does *not* make a copy:
``` {r }
obj2 <- obj
obj2[[2]] <- obj2[[2]] * 2
obj[[2]] == obj2[[2]]
```

For a better version of this, see
[storr](https://github.com/richfitz/storr) which does similar things to implement "[indexable serialisation](http://htmlpreview.github.io/?https://raw.githubusercontent.com/richfitz/storr/master/inst/doc/storr.html#lists-and-indexable-serialisation)"

What would be nice is a set of tools for working with any R/`Redis`
package that can convert R objects into `Redis` data structures so
that they can be accessed in pieces.  Of course, these objects
could be read/written by programs *other* than R if they also
support `Redis`.  We have made some approaches towards that with
the [docdbi](https://github.com/ropensci/docdbi) package, but this
is a work in progress.

# Getting help

Because `RedisAPI` package is simply a wrapper around the Redis
API, the main source of documentation is the Redis help itself,
`http://redis.io` (Unfortunately, the CRAN submission system
incorrectly complains and thinks that website is a dead link so I
direct links can't be included).  The Redis documentation is very
readable, thorough and contains great examples.

`RedisAPI` tries to bridge to this help.  Redis' commands are
"grouped" by data types or operation type; use
\code{redis_commands_groups()} to see these groups:
``` {r }
RedisAPI::redis_commands_groups()
```

To see command listed within a group, use the `redis_commands`
function:
``` {r }
RedisAPI::redis_commands("string")
```

Then use the function `redis_help` to get help for a particular
command:
``` {r }
RedisAPI::redis_help("SET")
```

The function definition here is the definition of the method you
will use within the retuned object (see below).  A default argument
of `NULL` indicates that a command is optional (`EX`, `PX` and
`condition` here are all optional).  The sentence is straight from
the Redis documentation, and the link will take you to the full
documentation on the Redis site.
